### 2020-04-16 paper/article.tex
more

% The MIT License (MIT)
%
% Copyright (c) 2020 Yegor Bugayenko
%
% Permission is hereby granted, free of charge, to any person obtaining a copy
% of this software and associated documentation files (the "Software"), to deal
% in the Software without restriction, including without limitation the rights
% to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
% copies of the Software, and to permit persons to whom the Software is
% furnished to do so, subject to the following conditions:
%
% The above copyright notice and this permission notice shall be included
% in all copies or substantial portions of the Software.
%
% THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
% IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
% FITNESS FOR A PARTICULAR PURPOSE AND NON-INFRINGEMENT. IN NO EVENT SHALL THE
% AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
% LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
% OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
% SOFTWARE.

% \documentclass[a4paper,UKenglish,cleveref, autoref]{lipics-v2019}
\documentclass[12pt]{article}
\title{Source Code Volatility Metric}
\author{Yegor Bugayenko}{}{}
% \authorrunning{Yegor Bugayenko}
% \ccsdesc[100]{Software and its engineering~Object oriented languages}
% \keywords{Object-Oriented Programming, Immutability, NCSS}
\usepackage[utf8]{inputenc}
\usepackage[numbers]{natbib}
\bibliographystyle{plainnat}
\usepackage{textcomp}
\usepackage[inline]{enumitem}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{verbatimbox}
\usepackage{interval}
\usepackage{hyperref}
\usepackage{minted}
  \setminted{fontsize=\footnotesize}
  \setminted{breaklines}
  \usemintedstyle{bw}
\newcommand{\code}[1]{\texttt{#1}}
\newenvironment{nicetable}
  {\setlength{\parindent}{0em}\medskip\small}
  {\medskip}
\input{total}
\begin{document}
\raggedbottom
\maketitle

\begin{abstract}
A new metric was introduced to calculate the distance
between actively modified files in a source code repository
and the files, which are rarely modified and may be considered
as ``dead'' or ``abandoned.'' It was empirically demonstrated that larger repositories
have larger values of the introduced metric.
\end{abstract}

\section{Introduction}

Most software development projects keep their source code in Git,
which is the de-facto standard in the industry at the moment, or similar
version control systems. Every system, including open
source products like Git, Subversion, and Mercurial, and commerical tools
like Borland StarTeam\texttrademark{} or IBM ClearCase\texttrademark{}
have the same feature: keeping track of the changes
made to the source code files, known as ``logs.''

Git logs provide information about every single change made by every software
developer during the entire course of the project. Using this information
it's possible to measure which files are being modified frequently. On the
other hand, it's also possible to spot files that are rarely modified and may
be considered as ``dead'' or ``abandoned'' code, which is a threat
to the maintainability of the entire project.

It is possible to introduce a metric to measure the relationship
between the amount of actively modified files and files which stay
in the repository for a long time without modifications.
In Section~\ref{sec:method} such a metric is introduced.

In order to demonstrate how the metric works we applied it to
\thetotalrepos{} public Java repositories from GitHub and analyzed
its impact on a few other simple metrics, including the number
of files in a repository, the number of directories, and the amount
of bytes.

\section{Related Work}

...

\section{Method}
\label{sec:method}

First, by looking at the Git history,
it is observed how many times every source code file out of $N$ was touched
during the lifetime of the repository (excluding the files that don't exist
in the repository anymore):

\begin{eqnarray}
T = [t_1, t_2, \dots, t_N]
\end{eqnarray}

Then, the entire interval between $\check{T}$ (the maximum value)
and $\hat{T}$ (the minimum value) is divided to $Z$ equivalent groups:

\begin{align}
G &= [g_1, g_2, \dots, g_{Z}] \\
\delta &= ( \check{T} - \hat{T} ) / Z \\
g_j &= \sum_{i=1}^N [ j(\delta-1) < t_i < j\delta ]
\end{align}

Then, the mean $\mu$ is calculated as:

\begin{eqnarray}
\mu = \frac{1}{Z}\sum_{j=1}^{Z}{g_j}
\end{eqnarray}

Finally, the variance is calculated as:

\begin{eqnarray}
\text{Var}(g) = \frac{1}{Z}\sum_{j=1}^{Z}{|g_j - \mu|^2}
\end{eqnarray}

The variance $\text{Var}(g)$ is the volatility of the source code. The smaller
the volatility the more cohesive is the repository and the smaller
the amount of the dead code inside it.

\section{Empirical Results}

A list of Java repositories were retrieved from GitHub via their
public API. The first \thetotalrepos{} repositories were taken, which satisfied
the selection criteria:
\begin{enumerate*}[label={\arabic*)}]
\item more than 1,000 GitHub stars,
\item more than 200 Kb of data,
\item not archived, and
\item public.
\end{enumerate*}
The list included popular Java open source products, such as
Spring, RxJava, Guava, MyBatis, Clojure, JUnit, Lombok,
Graal, Selenium, Spark, Mockito, Neo4j, Jenkins, Netty, and others.

The volatility metric was calculated for each repository, using the
formula explained above (the value of $Z$ was set to 64).
Then, a few other metrics were collected
for each repository and their values were compared with the volatility.

The Figure~\ref{fig:1} demonstrates the relationship between
the number of files in the repository~($M_1$) and its volatility~($V$). Both
axixes of the graph have logarithmic scales, for the sake of visual
understandability: the difference between minimum and maximum values
of volatility is logarithmically large. It is visually obvious that
repositories with larger number of files tend to have higher values
of the volatility metric.

\begin{figure}[h]
  \input{files.tex}
  \caption{The relationship between the number of files in the repository and its volatility}
  \label{fig:1}
\end{figure}

The Figure~\ref{fig:2} demonstrates the relationship between
the logarithm of the size of the Git repository in bytes~($M_2$) and
the logarithm of its volatility~($V$).
It is visually obvious that
binary-size larger repositories tend to have higher values
of the volatility metric.

\begin{figure}[h]
  \input{bytes.tex}
  \caption{The relationship between the size of the Git repository in bytes and its volatility}
  \label{fig:2}
\end{figure}

The Figure~\ref{fig:3} demonstrates the relationship between
the logarithm of the size of the Git repository in bytes~($M_3$) and
the logarithm of its volatility~($V$).
It is visually obvious that
repositories with more directories tend to have higher values
of the volatility metric.

\begin{figure}[h]
  \input{dirs.tex}
  \caption{The relationship between the number of repositories in the repository and its volatility}
  \label{fig:3}
\end{figure}

\section{Conclusion}

A new source code volatility metric was introduced and applied
to \thetotalrepos{} open source Java repositories from GitHub. It was
empirically demonstrated that larger repositories have higher values
of the volatility metric.

The source code of Ruby and Python scripts used to do the research
is available in GitHub repository \texttt{yegor256/volatility-vs-size}.

\bibliography{main}

\end{document}

### 2020-04-16 draw.rb

@@ -1,0 +1,64 @@
+#!/usr/bin/env ruby
+# The MIT License (MIT)
+#
+# Copyright (c) 2020 Yegor Bugayenko
+#
+# Permission is hereby granted, free of charge, to any person obtaining a copy
+# of this software and associated documentation files (the "Software"), to deal
+# in the Software without restriction, including without limitation the rights
+# to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+# copies of the Software, and to permit persons to whom the Software is
+# furnished to do so, subject to the following conditions:
+#
+# The above copyright notice and this permission notice shall be included
+# in all copies or substantial portions of the Software.
+#
+# THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+# IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+# FITNESS FOR A PARTICULAR PURPOSE AND NON-INFRINGEMENT. IN NO EVENT SHALL THE
+# AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+# LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+# OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+# SOFTWARE.
+
+require 'slop'
+
+opts = Slop.parse(ARGV, strict: true, help: true) do |o|
+  o.integer '--width', default: 12
+  o.integer '--height', default: 8
+  o.string '--xlabel', default: 'log_{10}(M)'
+  o.string '--ylabel', default: 'log_{10}(V)'
+  o.string '--xaxis', default: 'summary-files.csv'
+  o.string '--yaxis', default: 'summary-scv.csv'
+end
+
+points = []
+xaxis = File.open(opts[:xaxis]).read.lines
+yaxis = File.open(opts[:yaxis]).read.lines
+raise 'Invalid size' if xaxis.length != yaxis.length
+(0..xaxis.length-1).each do |i|
+  points << { x: xaxis[i].to_f, y: yaxis[i].to_f }
+end
+
[…]

### 2020-04-16 paper/article.tex
cite

@@
 
-Most software development projects keep their source code in Git,
+Most software development projects keep their source code in Git~\citet{loeliger2012},
 which is the de-facto standard in the industry at the moment, or similar
@@
 of files in a repository, the number of directories, and the amount
-of bytes.
+of bytes. Since, according to~\citet{muthukumaran2015},
+``code change metrics mined from source control repositories have
+proven to be the most reliable predictors of bugs in
+contemporary software engineering research,'' the introduced volatility
+metric may be used to spot maintainability issues.
 

### 2020-04-17 paper/article.tex
cite

@@
 
-...
+There are a few existing code change metrics, including code churn~\citep{munson1998}
+and others~\citep{fentonsoftware}.
 
@@
 \begin{align}
+\delta &= ( \check{T} - \hat{T} ) / Z \\
 G &= [g_1, g_2, \dots, g_{Z}] \\
-\delta &= ( \check{T} - \hat{T} ) / Z \\
 g_j &= \sum_{i=1}^N [ j(\delta-1) < t_i < j\delta ]
@@
 
-The Figure~\ref{fig:3} demonstrates the relationship between
-the logarithm of the size of the Git repository in bytes~($M_3$) and
-the logarithm of its volatility~($V$).
-It is visually obvious that
-repositories with more directories tend to have higher values
-of the volatility metric.
+\section{Discussion}
 
-\begin{figure}[h]
-  \input{dirs.tex}
-  \caption{The relationship between the number of repositories in the repository and its volatility}
-  \label{fig:3}
-\end{figure}
+It was demonstrated that the size of the source code is an important
+factor negatively affecting maintainability and other quality characteristicts.
+For example, \citep{heitlager2007} suggested to consider ``volume'' (the amount of code
+in the repository) as a primary contributor to an aggregate maintainability index
+of a software because ``it is fairly intuitive that the total size of a system should
+feature heavily in any measure of maintainability.''
 

### 2020-04-17 paper/article.tex
cite

@@
 
-There are a few existing code change metrics, including code churn~\citep{munson1998}
-and others~\citep{fentonsoftware}.
+The metric introduced in this research belongs to the family of
+``code change metrics,'' which analyze the history of changes in a repository~\citep{fentonsoftware}.
+Code churn is one of the simplest metric, which counts
+all lines being modified (added, edited, and deleted) during
+the entire lifetime of a project~\citep{munson1998}.
+The number of changes being made to a particular file, class, method
+or line of code is yet another simple metric, used for example by~\citet{demeyer2000}
+in order to analyze the effect of refactorings.
+
+\citep{biazzini2014} introduced a number of metrics to analyze
+Git history (specifically related to GitHub), such as unique-count, unique-ratio,
+VIP-count, VIP-ratio, scattered-count, pervasive-count and a few others. Some
+of the metrics are supposed to be calculated for a repository together with
+its forks, while others may be calculated for a single repository history.
+
+To our knowledge, the metric introduced in this research has never been
+suggested.
 
@@
 factor negatively affecting maintainability and other quality characteristicts.
-For example, \citep{heitlager2007} suggested to consider ``volume'' (the amount of code
-in the repository) as a primary contributor to an aggregate maintainability index
+For example, \citet{heitlager2007} suggested to consider ``volume'' (the amount of code
+in the repository, the number of lines per file, the number of methods in a class,
+and so on) as a primary contributor to an aggregate maintainability index
 of a software because ``it is fairly intuitive that the total size of a system should

### 2020-04-17 paper/article.tex
cite

@@
 
-\citep{biazzini2014} introduced a number of metrics to analyze
+\citet{biazzini2014} introduced a number of metrics to analyze
 Git history (specifically related to GitHub), such as unique-count, unique-ratio,
@@
 its forks, while others may be calculated for a single repository history.
+
+\citet{batista2018} provided a detailed analysis of existing metrics
+and introduced a new one to measure, by looking at Git/GitHub project
+commit history, how ``close'' developers stay to each other and form
+pairs.
 

### 2020-04-17 paper/article.tex
cite

@@
 The metric introduced in this research belongs to the family of
-``code change metrics,'' which analyze the history of changes in a repository~\citep{fentonsoftware}.
+``code change metrics,'' which analyze the history of changes in a repository~\citep{fentonsoftware,choudhary2018}.
 Code churn is one of the simplest metric, which counts

### 2020-04-17 paper/article.tex
cite

@@
 all lines being modified (added, edited, and deleted) during
-the entire lifetime of a project~\citep{munson1998}.
+the entire lifetime of a project~\citep{munson1998}. It was demonstrated
+by~\citet{shin2010} how code churn, together with complexity and other
+developer activity metrics, is related to software vulnerabilities.
 The number of changes being made to a particular file, class, method
@@
 in order to analyze the effect of refactorings.
+
+\citet{moser2008} made a comparative analysis of 17 change metrics to understand
+their efficiency for defect prediction: number of distinct authors per file,
+total modifications per file, total additions per file, maximum number of
+files committed together, and others.
 

### 2020-04-19 paper/article.tex
more text

@@
 
-% \documentclass[a4paper,UKenglish,cleveref, autoref]{lipics-v2019}
-\documentclass[12pt]{article}
+\documentclass[sigconf]{acmart}
 \title{Source Code Volatility Metric}
 \author{Yegor Bugayenko}{}{}
-% \authorrunning{Yegor Bugayenko}
-% \ccsdesc[100]{Software and its engineering~Object oriented languages}
-% \keywords{Object-Oriented Programming, Immutability, NCSS}
+\ccsdesc[100]{Software and its engineering~Metrics}
+\keywords{Metrics, Software Size, Software Maintainability}
 \usepackage[utf8]{inputenc}
-\usepackage[numbers]{natbib}
-\bibliographystyle{plainnat}
 \usepackage{textcomp}
@@
 \begin{document}
-\raggedbottom
-\maketitle
-
 \begin{abstract}
@@
 have larger values of the introduced metric.
+The metric may be used for earlier detection of code maintenance anomalies
+and helping software developers make the decision of splitting the repository
+into smaller oners in order to prevent maintainability issues.
 \end{abstract}
+\maketitle
 
@@
 
+The paper is organized as follows.
+Section~\ref{sec:background} defines various terms used in the paper.
+Section~\ref{sec:related} covers related work in the areas of code change metrics
+and empirical analysis of source code repository size.
+Section~\ref{sec:method} covers the source code volatility metric.
+Section~\ref{sec:results} covers our empirical case study of \thetotalrepos{} Java source code repositories.
+Section~\ref{sec:discussion} covers limitations of both the metric and the study.
+Finally, we summarize our study in Section~\ref{sec:conclusion}.
+
+\section{Background}
+\label{sec:background}
+
+In this paper, we use several terms regarding version control systems and
+software metrics. The type of metric we are proposing belongs to the family
+of source code change metrics, which analyze the history
+of changes in a source code repository~\citep{fentonsoftware,choudhary2018}.
+A \emph{change} in a version control system, such as Git, is an individual
+modification to the source code file made by a software developer locally
+on his/her machine and then \emph{committed} to the repository.
+
+Each version control system, including Git, provides an instrument for
+retrieving the entire history of changes from a repository. We use
+\texttt{git log} for Git.
+
 \section{Related Work}
+\label{sec:related}
 
 The metric introduced in this research belongs to the family of
-``code change metrics,'' which analyze the history of changes in a repository~\citep{fentonsoftware,choudhary2018}.
-Code churn is one of the simplest metric, which counts
+``code change metrics,''
+Code churn is one of the simplest metric in the family, which counts
 all lines being modified (added, edited, and deleted) during
@@
 developer activity metrics, is related to software vulnerabilities.
-The number of changes being made to a particular file, class, method
+Yet another simple metrict from the family is
+the number of changes being made to a particular file, class, method
 or line of code is yet another simple metric, used for example by~\citet{demeyer2000}
@@
 \section{Empirical Results}
+\label{sec:results}
 
@@
 \section{Discussion}
+\label{sec:discussion}
 
@@
 \section{Conclusion}
+\label{sec:conclusion}
 
@@
 
-The source code of Ruby and Python scripts used to do the research
-is available in GitHub repository \texttt{yegor256/volatility-vs-size}.
+% The source code of Ruby and Python scripts used to do the research
+% is available in GitHub repository \texttt{yegor256/volatility-vs-size}.
 
+\bibliographystyle{ACM-Reference-Format}
 \bibliography{main}

### 2020-04-19 draw.rb

@@ -24,8 +24,8 @@
 require 'slop'
 
 opts = Slop.parse(ARGV, strict: true, help: true) do |o|
-  o.integer '--width', default: 12
-  o.integer '--height', default: 8
+  o.integer '--width', default: 9
+  o.integer '--height', default: 5
   o.string '--xlabel', default: 'log_{10}(M)'
   o.string '--ylabel', default: 'log_{10}(V)'
   o.string '--xaxis', default: 'summary-files.csv'
@@ -56,7 +56,7 @@ puts "axis lines=middle, xlabel={$#{opts[:xlabel]}$}, ylabel={$#{opts[:ylabel]}$
 puts "xmin=#{points.map { |p| p[:x] }.min}, xmax=#{xmax},"
 puts "ymin=#{points.map { |p| p[:y] }.min}, ymax=#{ymax},"
 puts "extra tick style={major grid style=black},grid=major]"
-puts "\\addplot [mark=*, only marks] coordinates {"
+puts "\\addplot [mark=*, only marks, mark size=0.8pt] coordinates {"
 points.each do |p|
   puts "(#{p[:x]},#{p[:y]})"
 end

### 2020-04-19 paper/article.tex
more

@@
 \author{Yegor Bugayenko}{}{}
+\email{yegor.bugayenko@huawei.com}
+\affiliation{%
+  \institution{Huawei Technologies Co., Ltd.}
+  \city{Moscow, Russia}
+}
 \ccsdesc[100]{Software and its engineering~Metrics}
 \keywords{Metrics, Software Size, Software Maintainability}
+\setcopyright{acmcopyright}
+\copyrightyear{2020}
+\acmYear{2020}
+
 \usepackage[utf8]{inputenc}
@@
 and helping software developers make the decision of splitting the repository
-into smaller oners in order to prevent maintainability issues.
+into smaller ones in order to prevent maintainability issues.
 \end{abstract}
@@
 \texttt{git log} for Git.
+
+GitHub\texttrademark{} (currently owned by Microsoft\texttrademark{}) is
+one of the largest platforms for \emph{open source} projects.
+GitHub provides free access to all public Git repositories, which makes it possible
+to analyze Git histories.
 

### 2020-04-19 paper/article.tex
more

@@
 \documentclass[sigconf]{acmart}
-\title{Source Code Volatility Metric}
+\title{Volatility Metric to Detect Anomalies in Source Code Repositories}
 \author{Yegor Bugayenko}{}{}
@@
 
-Most software development projects keep their source code in Git~\citet{loeliger2012},
+Most software development projects keep their source code in Git~\citep{loeliger2012},
 which is the de-facto standard in the industry at the moment, or similar
@@
 have the same feature: keeping track of the changes
-made to the source code files, known as ``logs.''
+made to the source code files, also known as ``logs.''
 
@@
 other hand, it's also possible to spot files that are rarely modified and may
-be considered as ``dead'' or ``abandoned'' code, which is a threat
+be considered as ``dead'' or ``abandoned'' code, which may be considered as a threat
 to the maintainability of the entire project.
 
-It is possible to introduce a metric to measure the relationship
+We introduce a metric to measure the relationship
 between the amount of actively modified files and files which stay
 in the repository for a long time without modifications.
-In Section~\ref{sec:method} such a metric is introduced.
 
@@
 \thetotalrepos{} public Java repositories from GitHub and analyzed
-its impact on a few other simple metrics, including the number
-of files in a repository, the number of directories, and the amount
-of bytes. Since, according to~\citet{muthukumaran2015},
-``code change metrics mined from source control repositories have
-proven to be the most reliable predictors of bugs in
-contemporary software engineering research,'' the introduced volatility
-metric may be used to spot maintainability issues.
+its impact on a two other characteristics of each repository:
+the number of files in a repository and the size of it in bytes.
+Larger repositories shown larger values of the volatility.
 
@@
 
-The metric introduced in this research belongs to the family of
-``code change metrics,''
+``Code change metrics mined from source control repositories have
+proven to be the most reliable predictors of bugs in
+contemporary software engineering research,'' says \citet{muthukumaran2015}.
+
 Code churn is one of the simplest metric in the family, which counts
@@
 
-\section{Method}
+\section{Metric}
 \label{sec:method}
@@
 Then, the entire interval between $\check{T}$ (the maximum value)
-and $\hat{T}$ (the minimum value) is divided to $Z$ equivalent groups:
+and $\hat{T}$ (the minimum value) is divided to $Z$ (the input parameter of the method)
+equivalent groups:
 

### 2020-04-19 paper/article.tex
more

@@
 the number of files in a repository and the size of it in bytes.
-Larger repositories shown larger values of the volatility.
+Larger repositories shown larger values of the metric.
 
@@
 software metrics. The type of metric we are proposing belongs to the family
-of source code change metrics, which analyze the history
+of \emph{source code change metrics}, which analyze the history
 of changes in a source code repository~\citep{fentonsoftware,choudhary2018}.
-A \emph{change} in a version control system, such as Git, is an individual
+A \emph{change} in a version control system, such as Git, is an atomic
 modification to the source code file made by a software developer locally
@@
 by~\citet{shin2010} how code churn, together with complexity and other
-developer activity metrics, is related to software vulnerabilities.
+developer activity metrics, is related to software vulnerabilities. There
+are modifications of code churn, for example taking into accout socio-technical
+aspect of the metric, as suggested by~\citet{meneely2012}.
+
 Yet another simple metrict from the family is
@@
 
-To our knowledge, the metric introduced in this research has never been
+To our knowledge, the metrics introduced in this research, in order
+to detect anomalies in source code repository maintenance, has never been
 suggested.

### 2020-04-19 paper/article.tex
more

@@
 
-Yet another simple metrict from the family is
+Yet another simple metric from the family is
 the number of changes being made to a particular file, class, method
-or line of code is yet another simple metric, used for example by~\citet{demeyer2000}
+or line of code, which is used for example by~\citet{demeyer2000}
 in order to analyze the effect of refactorings.
@@
 
-To our knowledge, the metrics introduced in this research, in order
+To our knowledge, the metrics introduced in this research in order
 to detect anomalies in source code repository maintenance, has never been
@@
 
+It would be beneficial in future researches to analyze the relationship between
+volatility and other parameters of source code repositories,
+such as
+the programming language being used,
+the popularity in GitHub (the number of stars and forks),
+the total number of lines of code,
+the average number of lines of code per file,
+the number of Git commit authors,
+the total age of the repository,
+the number of issues and pull requests in GitHub,
+and many others.
+Would be interesting to empiricially demonstrate which of these parameters
+(or combinations of them)
+impact the behavior of volatility and in which direction.
+
 \section{Conclusion}

### 2020-04-19 paper/article.tex
more

@@
 and the files, which are rarely modified and may be considered
-as ``dead'' or ``abandoned.'' It was empirically demonstrated that larger repositories
+abandoned or even dead. It was empirically demonstrated that larger repositories
 have larger values of the introduced metric.
@@
 other hand, it's also possible to spot files that are rarely modified and may
-be considered as ``dead'' or ``abandoned'' code, which may be considered as a threat
+be considered as abandoned code, which may be considered as a threat
 to the maintainability of the entire project.
@@
 the volatility the more cohesive is the repository and the smaller
-the amount of the dead code inside it.
+the amount of the abandonded code inside it.
 
@@
 axixes of the graph have logarithmic scales, for the sake of visual
-understandability: the difference between minimum and maximum values
-of volatility is logarithmically large. It is visually obvious that
+understandability: the difference between the minimum and the maximum values
+of the volatility is logarithmically large. It is visually obvious that
 repositories with larger number of files tend to have higher values
@@
 \begin{figure}[h]
-  \input{files.tex}
-  \caption{The relationship between the number of files in the repository and its volatility}
+  \input{files-64.tex}
+  \caption{The relationship between the number of files in the repository and its volatility ($Z$ is set to 64)}
   \label{fig:1}
@@
 \begin{figure}[h]
-  \input{bytes.tex}
-  \caption{The relationship between the size of the Git repository in bytes and its volatility}
+  \input{bytes-64.tex}
+  \caption{The relationship between the size of the Git repository in bytes and its volatility ($Z$ is set to 64)}
   \label{fig:2}
+\end{figure}
+
+The Figure~\ref{fig:1} demonstrates the same relationship as the Figure~\ref{fig:1},
+but the parameter of the volatility formula is set to 32 instead of 64. The
+trend of the graph didn't change much.
+
+\begin{figure}[h]
+  \input{files-32.tex}
+  \caption{The relationship between the number of files in the repository and its volatility ($Z$ is set to 32)}
+  \label{fig:3}
+\end{figure}
+
+The Figure~\ref{fig:1} demonstrates the same relationship as the Figure~\ref{fig:1},
+but this time the parameter of the volatility formula is increased to 128
+instead of the original value of 64. The trend of the graph didn't change much.
+
+\begin{figure}[h]
+  \input{files-128.tex}
+  \caption{The relationship between the number of files in the repository and its volatility ($Z$ is set to 128)}
+  \label{fig:4}
 \end{figure}
@@
 factor negatively affecting maintainability and other quality characteristicts.
-For example, \citet{heitlager2007} suggested to consider ``volume'' (the amount of code
-in the repository, the number of lines per file, the number of methods in a class,
+For example, \citet{heitlager2007} suggested to consider ``volume''
+(the number of lines per file, the number of methods in a class,
 and so on) as a primary contributor to an aggregate maintainability index
@@
 feature heavily in any measure of maintainability.''
+
+However, the discussion of whether larger repositories are a better
+practice than smaller ones is still open. Google, for example, advocates
+for larger ones, known as ``monolithic,'' which tend to keep many modules
+and components in one Git space, simplifying dependencies management~\citep{jaspan2018}.
+As was demonstrated in our research, larger repositories will tend to
+have maintenance anomalies: some parts of source code will be changed much
+less frequently than other parts. Moreover, the distance between ``popular''
+parts and ``abandoned'' ones will grow when the size of the repository grows.
+
+These maintenance anomalies will lead to maintability and quality problems
+mostly because it will be difficult for programmers to quickly switch between
+different contexts inside one repository: one set of files is being modified
+every day, while another one stays untouched for months. A programmer will
+try to avoid making changes to the part that is abandonded because it won't
+look like code that is expecting changes.
+
+Also, maintenance anomalies is a sign of differences in customer demand
+for the source code components. The module with smaller frequency of changes
+must be released less frequently than the module that is being edited every day.
+Not paying attention to the growing volatility of the repository means
+ignoring anomalies and, because of that, redundant releasing of the code
+that doesn't require new releases. Moreover, releasing larger software modules
+usually means longer build cycles, which are the key failure factor of
+continuous delivery, as explained by~\cite{humble2010}. Getting rid of
+maintenance anomalies may help reduce build cycles.
 

### 2020-05-17 paper/article.tex
more text

@@
 
-\documentclass[sigconf]{acmart}
+\documentclass[anonymous,sigconf,10pt,nonacm=true]{acmart}
 \title{Volatility Metric to Detect Anomalies in Source Code Repositories}
@@
 \keywords{Metrics, Software Size, Software Maintainability}
-\setcopyright{acmcopyright}
-\copyrightyear{2020}
-\acmYear{2020}
+% \setcopyright{acmcopyright}
+% \copyrightyear{2020}
+% \acmYear{2020}
 
@@
 developer during the entire course of the project. Using this information
-it's possible to measure which files are being modified frequently. On the
-other hand, it's also possible to spot files that are rarely modified and may
-be considered as abandoned code, which may be considered as a threat
-to the maintainability of the entire project.
+it's possible to measure which files are being modified more frequently than others.
+On the other hand, it's also possible to spot files that are rarely modified and may
+be considered either as \emph{abandoned} or \emph{stable} code.
+\citet{fontana2014} claims that the gap between actively modified code
+and the code that is stable should be as narrow as possible in a properly
+maintained repository.
 
-We introduce a metric to measure the relationship
+Indeed, abandoned code may be considered as a threat
+to the maintainability of the entire project, because its lifecycle
+is not aligned with the lifecycle of active code, which is modified more intensively.
+Later, when a modification is required to the abandonded code, a programmer
+may have difficulties finding out the principles of design of the code and
+making sure new modifications don't break it.
+
+We introduce a Volatility metric to measure the relationship
 between the amount of actively modified files and files which stay
 in the repository for a long time without modifications.
+The hipothesis is that the Volatility is related to the size of the
+repository: the larger the repository the higher the volatility. Larger
+repositories are an obvious threat to code maintainability. Thus, if
+the relationship between size and Volatility exists, the new metric
+can be used to identify code anomalies and detect badly maintained repositories.
 
-In order to demonstrate how the metric works we applied it to
+In order to validate the hipothesis and demonstrate how the metric works we applied it to
 \thetotalrepos{} public Java repositories from GitHub and analyzed
-its impact on a two other characteristics of each repository:
+its impact on two other characteristics of each repository:
 the number of files in a repository and the size of it in bytes.
+
 Larger repositories shown larger values of the metric.
@@
 
-To our knowledge, the metrics introduced in this research in order
+\citet{fontana2014} suggested to measure ``File Volatility'' as the ratio
+between the maximum number of changes happened to a file line and the number
+of all changes to the file since its creation. Going further, they also suggested to
+calculate ``Repository Stability'' metric as a relationship between
+active files (with high File Volatility) and closed files.
+The also concluded that repositories where most files change
+over time can suffer from organizational or design issues.
+
+To our knowledge, the metric introduced in this research, in order
 to detect anomalies in source code repository maintenance, has never been
@@
 
-Then, the entire interval between $\check{T}$ (the maximum value)
+Here, $t_i$ is the number of commits found in the repository, which
+modified the $i$-th file. Then, the entire interval between $\check{T}$ (the maximum value)
 and $\hat{T}$ (the minimum value) is divided to $Z$ (the input parameter of the method)
@@
 G &= [g_1, g_2, \dots, g_{Z}] \\
-g_j &= \sum_{i=1}^N [ j(\delta-1) < t_i < j\delta ]
+g_j &= \sum_{i=1}^N [ j \times (\delta-1) < t_i < j \times \delta ]
 \end{align}
+
+Here, $g_j$ is the total number of $t_i$, which are larger than
+$j \times (\delta-1)$ and smaller than $j \times \delta$. In other words,
+$T$, the entire set of measurements, is sorted and then split into $Z$ sectors,
+where $g_j$ is the counter of measurements that belong to the sector $j$.
 
@@
 
-The variance $\text{Var}(g)$ is the volatility of the source code. The smaller
-the volatility the more cohesive is the repository and the smaller
-the amount of the abandonded code inside it.
+The variance $\text{Var}(g)$ is the Volatility of the source code. The smaller
+the Volatility the more cohesive is the repository and the smaller
+the amount of the abandonded code inside it. The metric demonstrates how
+large is the gap between intensively modified files and the abandoned ones:
+the larger the variance, the wider the graph, and the larger the gap.
+A repository with a thousand files all being modified at the same pace would
+have a Volatility of zero. On the other hand, a repository with just a few
+files, where one of them is modified a thousand times, while others are modified
+just once, will have the Volatility of a miillion.
 
@@
 
-The volatility metric was calculated for each repository, using the
+The Volatility metric was calculated for each repository, using the
 formula explained above (the value of $Z$ was set to 64).
 Then, a few other metrics were collected
-for each repository and their values were compared with the volatility.
+for each repository and their values were compared with the Volatility.
 
 The Figure~\ref{fig:1} demonstrates the relationship between
-the number of files in the repository~($M_1$) and its volatility~($V$). Both
-axixes of the graph have logarithmic scales, for the sake of visual
+the number of files in the repository~($M_1$) and its Volatility~($V$).
+Both axixes of the graph have logarithmic scales, for the sake of visual
 understandability: the difference between the minimum and the maximum values
-of the volatility is logarithmically large. It is visually obvious that
+of the Volatility is logarithmically large. It is visually obvious that
 repositories with larger number of files tend to have higher values
-of the volatility metric.
+of the Volatility metric.
 
@@
   \input{files-64.tex}
-  \caption{The relationship between the number of files in the repository and its volatility ($Z$ is set to 64)}
+  \caption{The relationship between the number of files in the repository and its Volatility ($Z$ is set to 64)}
   \label{fig:1}
@@
 the logarithm of the size of the Git repository in bytes~($M_2$) and
-the logarithm of its volatility~($V$).
+the logarithm of its Volatility~($V$).
 It is visually obvious that
 binary-size larger repositories tend to have higher values
-of the volatility metric.
+of the Volatility metric.
 
@@
   \input{bytes-64.tex}
-  \caption{The relationship between the size of the Git repository in bytes and its volatility ($Z$ is set to 64)}
+  \caption{The relationship between the size of the Git repository in bytes and its Volatility ($Z$ is set to 64)}
   \label{fig:2}
@@
 The Figure~\ref{fig:1} demonstrates the same relationship as the Figure~\ref{fig:1},
-but the parameter of the volatility formula is set to 32 instead of 64. The
+but the parameter of the Volatility formula is set to 32 instead of 64. The
 trend of the graph didn't change much.
@@
   \input{files-32.tex}
-  \caption{The relationship between the number of files in the repository and its volatility ($Z$ is set to 32)}
+  \caption{The relationship between the number of files in the repository and its Volatility ($Z$ is set to 32)}
   \label{fig:3}
@@
 The Figure~\ref{fig:1} demonstrates the same relationship as the Figure~\ref{fig:1},
-but this time the parameter of the volatility formula is increased to 128
+but this time the parameter of the Volatility formula is increased to 128
 instead of the original value of 64. The trend of the graph didn't change much.
@@
   \input{files-128.tex}
-  \caption{The relationship between the number of files in the repository and its volatility ($Z$ is set to 128)}
+  \caption{The relationship between the number of files in the repository and its Volatility ($Z$ is set to 128)}
   \label{fig:4}
@@
 must be released less frequently than the module that is being edited every day.
-Not paying attention to the growing volatility of the repository means
+Not paying attention to the growing Volatility of the repository means
 ignoring anomalies and, because of that, redundant releasing of the code
@@
 It would be beneficial in future researches to analyze the relationship between
-volatility and other parameters of source code repositories,
+Volatility and other parameters of source code repositories,
 such as
@@
 (or combinations of them)
-impact the behavior of volatility and in which direction.
+impact the behavior of Volatility and in which direction.
 
@@
 
-A new source code volatility metric was introduced and applied
+A new source code Volatility metric was introduced and applied
 to \thetotalrepos{} open source Java repositories from GitHub. It was
 empirically demonstrated that larger repositories have higher values
-of the volatility metric.
+of the Volatility metric.
 

### 2021-09-13 paper/article.tex
edited

@@
 
-\documentclass[anonymous,sigconf,10pt,nonacm=true]{acmart}
+\documentclass[sigplan,10pt,nonacm=true]{acmart}
+\settopmatter{printfolios=false,printccs=false,printacmref=false}
+
 \title{Volatility Metric to Detect Anomalies in Source Code Repositories}
 \author{Yegor Bugayenko}{}{}
-\email{yegor.bugayenko@huawei.com}
-\affiliation{%
-  \institution{Huawei Technologies Co., Ltd.}
-  \city{Moscow, Russia}
-}
+\email{yegor256@gmail.com}
+\affiliation{\institution{}\city{Moscow}\country{Russia}}
 \ccsdesc[100]{Software and its engineering~Metrics}
@@
   {\medskip}
-\input{total}
+\input{tex/total}
 \begin{document}
@@
 version control systems. Every system, including open
-source products like Git, Subversion, and Mercurial, and commerical tools
-like Borland StarTeam\texttrademark{} or IBM ClearCase\texttrademark{}
+source products like Git, Subversion, and Mercurial, and commercical tools
+like StarTeam\texttrademark{} or ClearCase\texttrademark{}
 have the same feature: keeping track of the changes
@@
 \begin{figure}[h]
-  \input{files-64.tex}
+  \input{tex/files-64.tex}
   \caption{The relationship between the number of files in the repository and its Volatility ($Z$ is set to 64)}
@@
 \begin{figure}[h]
-  \input{bytes-64.tex}
+  \input{tex/bytes-64.tex}
   \caption{The relationship between the size of the Git repository in bytes and its Volatility ($Z$ is set to 64)}
@@
 \begin{figure}[h]
-  \input{files-32.tex}
+  \input{tex/files-32.tex}
   \caption{The relationship between the number of files in the repository and its Volatility ($Z$ is set to 32)}
@@
 \begin{figure}[h]
-  \input{files-128.tex}
+  \input{tex/files-128.tex}
   \caption{The relationship between the number of files in the repository and its Volatility ($Z$ is set to 128)}

### 2021-09-18 paper/article.tex
sent to bcnc

@@
 
-\documentclass[sigplan,10pt,nonacm=true]{acmart}
-\settopmatter{printfolios=false,printccs=false,printacmref=false}
+\documentclass[sigplan,10pt,screen]{acmart}
+% \settopmatter{printfolios=false,printccs=false,printacmref=false}
 
 \title{Volatility Metric to Detect Anomalies in Source Code Repositories}
-\author{Yegor Bugayenko}{}{}
+\author{Yegor Bugayenko}
+\orcid{0000-0001-6370-0678}
 \email{yegor256@gmail.com}
 \affiliation{\institution{}\city{Moscow}\country{Russia}}
-\ccsdesc[100]{Software and its engineering~Metrics}
+
+\setcopyright{acmcopyright}
+\acmPrice{15.00}
+\acmDOI{10.1145/3486949.3486961}
+\acmYear{2021}
+\copyrightyear{2021}
+\acmSubmissionID{splashws21bcncmain-p41-p}
+\acmISBN{978-1-4503-9125-2/21/10}
+\acmConference[BCNC '21]{Proceedings of the 1st ACM SIGPLAN International Workshop on Beyond Code: No Code}{October 17, 2021}{Chicago, IL, USA}
+\acmBooktitle{Proceedings of the 1st ACM SIGPLAN International Workshop on Beyond Code: No Code (BCNC '21), October 17, 2021, Chicago, IL, USA}
+
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10011007.10011074.10011111.10011696</concept_id>
+<concept_desc>Software and its engineering~Maintaining software</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+\ccsdesc[500]{Software and its engineering~Maintaining software}
 \keywords{Metrics, Software Size, Software Maintainability}
-% \setcopyright{acmcopyright}
-% \copyrightyear{2020}
-% \acmYear{2020}
 
@@
 
-% The source code of Ruby and Python scripts used to do the research
-% is available in GitHub repository \texttt{yegor256/volatility-vs-size}.
+The source code of Ruby and Python scripts used to do the research
+is available in GitHub repository:\newline\texttt{yegor256/volatility-vs-size}.
 
@@
 \bibliography{main}
+\clearpage
 
